---
title: "p8105_hw3_yx2711"
author: "Yingchen Xu"
date: "2022-10-15"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
library(ggridges)
library(patchwork)
library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


# Problem 1

Load dataset `instacart`
```{r}
data("instacart")

instacart = instacart %>%
  as_tibble(instacart) %>% 
  janitor::clean_names()
```

There are `r nrow(instacart)` rows and `r ncol(instacart)` columns in the `instacart` dataset, with each row representing one product of the order. Key variables include order details (order number, order time, etc), user id and product detail (product id, product name, products' departments). In total, there are `r instacart %>% select(product_id) %>% distinct %>% count` products found in `r instacart %>% select(user_id, order_id) %>% distinct %>% count` orders from `r instacart %>% select(user_id) %>% distinct %>% count` distinct users.


Count number of aisles and identify which aisles are the most items ordered from.
```{r}
instacart %>% 
  select(aisle) %>% 
  distinct %>% 
  count()

instacart %>% 
  group_by(aisle) %>% 
  summarize(obs = n()) %>% 
  arrange(desc(obs))
```

There are `r instacart %>% select(aisle) %>% distinct %>% count()` aisles and fresh vegetables and fresh fruits have the most orders. 


Make a plot that shows the number of items ordered in each aisle, arrange and organize.
```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(aisle = fct_reorder(aisle, n)) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```


Make a table showing the three most popular items in each of the aisles "baking ingredients", "dog food care', and "packaged vegetables fruit."
```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(desc(n)) %>%
  knitr::kable()
```


Make a table showing the mean hour the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week
```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  spread(key = order_dow, value = mean_hour) %>%
  knitr::kable(digits = 2)
```


# Problem 2

Read file and clean names.
Use `mutate()` to add a type variable to indicate that whether the day is a weekend or a weekday.
Use `Pivot_longer()` to transform the dataset from the wide format to a long format and name the columns as "activity_id" and "counts"
```{r}
activity = read.csv("data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    type = ifelse(day == c("Saturday", "Sunday"), "weekend", "weekday")
  ) %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to = "minute",
    names_prefix = "activity_",
    values_to = "counts"
  )
```

The dataset `activity` contains `r nrow(activity)` observations and `r ncol(activity)` variables. The variables are week, day_id, day, type (indicating whether it is a weekday or weekend), minute, and activity counts. 


Use `knitr::kable()` to create a table showing the total activity for each day.
Also create a scatterplot to see whether there is a trend.
```{r}
activity %>% 
  group_by(day_id) %>% 
  summarize(sum_counts = sum(counts, na.rm = TRUE)) %>% 
  knitr::kable(digits = 2)

plot_total = activity %>% 
  group_by(day_id) %>% 
  summarize(sum_counts = sum(counts)) %>%
  ggplot(aes(x = day_id, y = sum_counts)) +
  geom_point() +
  geom_line() +
  labs(
    title = "The total activity over the day",
    x = "Day id",
    y = "Total activity counts",
    caption = "Data from the accel_data")

plot_total
```

By observing the table and the plot, it seems that there is no apparent trend between total activity and day_id. 


Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week.
Use `geom_line()` to show the overall trend and use `geom_smooth()` to indicate variations within each day of a week.
```{r}
plot_each_day = activity %>% 
  mutate(hour = as.numeric(minute) / 60) %>% 
  ggplot(aes(x = hour, y = counts, color = day)) +
  geom_line(alpha = 0.5) +
  geom_smooth(se = FALSE) +
  labs(
    title = "The 24-hour activity time courses for each day",
    x = "Hour",
    y = "Activity counts",
    caption = "Data from the accel_data") +
  scale_x_continuous(breaks = c(0, 4, 8, 12, 16, 20, 24))

plot_each_day
```

Based on the overall graph, we can observe that the overall activity counts are low from 0 am to around 5 am. The activity counts start to increase from 5 am to 9 am and has a peak around 7 am. Then the activity counts stay high and constant from 10 am to around 11 pm. 
Based on the `geom_smooth()`, we can observe there are differences between each day of a week. There is a distinct peak at around 10 am for Sunday and there is a distinct peak at around 10 pm for Friday.  


# Problem 3

Load dataset `ny_noaa`.
```{r}
data("ny_noaa")

ny_noaa = ny_noaa %>% 
  as_tibble(ny_noaa) %>% 
  janitor::clean_names()

```

The `ny_noaa` dataset contains `r nrow(ny_noaa)` rows and `r ncol(ny_noaa)` columns. The key variables are weather station id, date, precipitation, snowfall, snow depth, maximum temperature and minimum temperature. There are `r sum(is.na(ny_noaa$prcp))` missing values in the precipitation variables, `r sum(is.na(ny_noaa$snow))` missing values in the snow fall variables,
`r sum(is.na(ny_noaa$snwd))` missing values in the snow depth variables, `r sum(is.na(ny_noaa$tmax))` missing values in the maximum temperature variables and `r sum(is.na(ny_noaa$tmin))` missing values in the minimum temperature variables.


Create separate variables for year, month, and day and ensure observations for temperature, precipitation, and snowfall are given in reasonable units.
```{r}
ny_noaa = ny_noaa %>% 
  separate(date, c("year", "month", "day")) %>% 
  mutate(
    prcp = prcp / 10,
    tmax = as.numeric(tmax) / 10,
    tmin = as.numeric(tmin) / 10,
    year = as.numeric(year),
    month = as.numeric(month),
    day = as.numeric(day)
  )

ny_noaa %>% 
  group_by(snow) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
```

The date variable has been separated into three variables year, month and day. The units of precipitation has changed to mm and the units of tmax and tmin has changed to degrees C. 
The most commonly observed value for snowfall is 0. Since snowfall only occurs in the winter and for the rest of the year, the occurrence of snowfall would be 0. 


Make a two-panel plot showing the average max temperature in January and in July in each station across years.
```{r}
plot_Jan_vs_Jul = ny_noaa %>% 
  filter(
    month == c(1, 7),
    !is.na(tmax)) %>% 
  mutate(
    month = recode(month, "01" = "January", "07" = "July")
  ) %>% 
  group_by(id, month, year) %>% 
  summarize(avg_max = mean(tmax)) %>% 
  ggplot(aes(x = year, y = avg_max)) +
  geom_point() +
  geom_line() +
  facet_grid(.~month) +
  labs(
    title = "The average max temperature in January and July in each station across years",
    x = "Year",
    y = "Average maximum temerature (degrees C)",
    caption = "Data from p8105 package - `NY_NOAA`"
  ) + 
  scale_x_continuous(n.break = 10)
  

plot_Jan_vs_Jul
```

The overall average maximum temperature is higher in July in each station across years than in January. There are fluctuations in average maximum temperature in each station over the years in these two months. In each year, the ranges of average maximum temperatures in each station are smaller in July than in January on average. There are outliers. For example, in January 2004, there is one station that has average maximum temperature of 14 degrees C, which is much higher than the rest in January. In July 1988, there is one station that has average maximum temperature of around 14 degrees C, which is much lower than the rest in July. 


MAke a two-panel plot showing
(i) tmax vs tmin for the full data set
(ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year
Combine panels using tools in the `patchwork` package
```{r}
#plot (i)
plot_tmax_vs_tmin = ny_noaa %>% 
  filter(!is.na(tmax), !is.na(tmin)) %>% 
  ggplot(aes(x = tmax, y = tmin)) +
  geom_hex() +
  labs(
    title = "Maximum temperature vs \nMinimum temperature (degrees C)",
    x = "Maximum temperature (degrees C)",
    y = "Minimum temperature (degrees C)",
    caption = "Data from p8105 package - `NY_NOAA`") +
  scale_fill_continuous(breaks = c(20000, 40000))

plot_tmax_vs_tmin


#plot (ii)
plot_snowfall = ny_noaa %>% 
  filter(snow %in% 1:99) %>% 
  ggplot(aes(x = snow, y = as.factor(year), fill = year)) +
  geom_density_ridges(alpha = .5) +
  scale_x_continuous(breaks = c(0, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100)) +
  labs(
    title = "The distribution of snowfall \nseparately by year",
    x = "Snowfall (mm)",
    y = "Year",
    caption = "Data from p8105 package - `NY_NOAA`"
  )

plot_snowfall


# Combine panels using patchwork
plot_tmax_vs_tmin + plot_snowfall
           
```

By observing the tmax vs tmin for the full dataset, most of the maximum temperature and minimum temperature range from -10 to 30 degrees C. Higher amount of counts with maximum temperature around 25 degrees C and with minimum temperature around 15 degrees C. There are outliers and and some of the outliers have extreme maximum temperature and low to normal minimum temperature.

By observing the distribution of snowfall separately by year, we can observe that there are some constant peaks of snowfall for each year, that are around 10mm, 25mm, 50mm, and 75mm. The peaks of snowfall is smaller for higher amount of snowfall. For example, the occurrence of snowfall of 75mm is less frequent than the occurrence of 10mm. The occurrence of high amount of snowfall is less frequent over the time, that the peaks at 75mm are smaller for year 2001 compared to year 1981. 

